---
title: 'Lab 02: R fundamentals 2'
author: "JT Larkin"
subtitle: "Partners: Ethan Rutledge, Abigail Guinan, and Sonja Glasser"
date: "9/15/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
n = 12345
vec_1 = sample(12, n, replace = TRUE)
head(vec_1)

vec_2 = vec_1 == 3

  
  
vec_1[vec_2]
```



1. Show the R code you used to create vec_2. :::{.warn} Your code should be a complete and self-contained example. I should be able to paste your code into a fresh R session on my computer and re-create your vec_2 :::

  * vec_2 = vec_1 == 3

2. Give two reasons why determining which elements in vec_1 have value 3 by visual inspection is a bad idea? 
  * There are 12,345 numbers in the vector and this would take a very long time and it is prone to error. 




```{r}
n = 12345
vec_1 = sample(12, n, replace = TRUE)
head(vec_1)


length(vec_1)


sum(vec_1 == 3)


n = 10
vec_1 = sample(12, n, replace = TRUE)
paste0("Sum of elements with value 3: ", sum(vec_1 == 3))

```



3. Why didn’t you always get the same count of 3 entries each time?
  * We did not always get the same count of 3 because the sample function is selecting 10 random numbers between 1 and 12. Each iteration R created a new set of 10 random numbers between 1 and 12.


4. Considering the different vectors generated each time, explain why using a logical test is a safe way to select entries with a value of 3.
  * It is the best method for selecting a specific entries. It cuts down on error and it is much more efficient. When we start to work with larger datasets this will especially be true. 
  

5. Explain why performing logical 'by hand' subsetting is very very bad practice. You may want consider re-usability of code, working with different sized data sets, and sharing code with collaborators.
  * This is a bad practice because there is a high likelihood of human error. This is especially true for large data sets.


```{r}
for (i in 1:10)
{
  print(paste0("This is loop iteration:", i))
}

```

6. Provide the code for your modified loop. It must run as a self-contained example on a fresh R session on my computer.
  * "for (i in 1:10)
{
  print(paste0("This is loop iteration:", i))
}"



```{r}
n = 16

for (i in 1:n)
{
  print(paste0("This is loop iteration:", i))
}
```



7. Provide the code for the modified loop that executes n times. It needs to be a self contained example. I should be able to set the value of n and then run your loop on my computer.
  * "n = 16

for (i in 1:n)
{
  print(paste0("This is loop iteration:", i))
}"


```{r}
n = 17

vec_1 = sample(10, n, replace = TRUE)

vec_1


for (i in 1:n)
{
  print(paste0("The element of vec_1 at index ", i, " is: ", vec_1[i]))
}
```



8. Provide the code you used to create the n, vec_1, and the loop. As always, it should run as a stand-alone example in a fresh R session on my computer.
  * "n = 17

vec_1 = sample(10, n, replace = TRUE)

vec_1


for (i in 1:n)
{
  print(paste0("The element of vec_1 at index ", i, " is: ", vec_1[i]))
}"



```{r}
create_and_print_vec = function(n, min = 1 , max = 10)
{
  vec_3 = sample(min:max, n, replace = TRUE)
  
  
  for (i in 1:n)
  {
  
  print(paste0("The element at index ", i, " is: ", vec_3[i]))
  }
}

create_and_print_vec(10)
create_and_print_vec(10, min = 100, max = 2000)

```

9. Provide the code you used to build your function.
  * To receive full credit your code must run without error on a new R session and produce output similar to the examples given in the instructions.
  
    * "create_and_print_vec = function(n, min = 1 , max = 10)
{
  vec_3 = sample(min:max, n, replace = TRUE)
  
  
  for (i in 1:n)
  {
  
  print(paste0("The element at index ", i, " is: ", vec_3[i]))
  }
}"