---
title: "Lab 07_The Bootstrap"
author: "JT Larkin (Partners: Ethan Rutledge and Sonja Glasser)"
date: "10/26/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
require(here)

### APPLY() FUNCTION

# Create simulated data
dat = matrix(1:49, nrow = 7, ncol = 7)
print(dat)

# Min and Max values in each row
apply(dat, MARGIN = 1, FUN = min)  # Margin = 1 for rows
apply(dat, MARGIN = 1, FUN = max)

# Mean values in each column
apply(dat, MARGIN = 2, FUN = mean) # Margin = 2 for columns

### DATA FILES
moths = read.csv(here("data", "moths.csv"))
head(moths)
```

```{r include = FALSE}
### MOTH ABUNDANCE: CIs
mean(moths$anst)
sd(moths$anst) # sd = 3.633761
moths$anst #sample size = 24

anst_SE = 3.633761/sqrt(24)

# Calculating critical t-values
# Function: qt(alpha value, df)
anst_crit_ts =  c(qt(0.05/2, 23), qt(1 - 0.05/2, 23)) # df = 23
anst_crit_ts

#Just the positive Critical T
anst_crit_t = qt(1 - 0.05/2, 23)
anst_crit_t

#CI radius (CI = mean +/- CI radius)

anst_CI_radius = anst_crit_t * anst_SE
anst_CI_radius

anst_CIs = c(mean(moths$anst) - anst_CI_radius,  mean(moths$anst) + anst_CI_radius)
anst_CIs
```

```{r include =FALSE}
### A SIMPLE BOOTSTRAP CI

# Create a results vector
m = 10000

# numeric() creates an vector of length m with all values initialized to zero
result = numeric(m)
head(result)

# Perform the bootstrap

for(i in 1:m)
{
  result[i] = mean(sample(moths$anst, replace=TRUE))
}

# Calculate the quantiles
mean(result)
quantile(result, c(0.025, 0.975))

### BOOTSTRAP INTERVAL USING boot() function

#install.packages("boot")
require(boot)
#boot(data, statistic, R)

# Custom mean function
boot_mean = function(x, i)
{
  return(mean(x[i], na.rm = TRUE))
}


myboot = 
  boot(
    data = moths$anst,
    statistic = boot_mean,
    R = 10000)
print(myboot)

str(myboot)

mean(moths$anst)
myboot$t0
mean(myboot$t) - myboot$t0
sd(myboot$t)

# Bootstrap CIs
quantile(
  myboot$t,
  c(0.025, 0.975))


```

```{r include = FALSE, eval=FALSE}
### RAREFACTION CURVE

#Setting up the bootstrap
moth_dat = moths[,-1]
head(moth_dat)

n = nrow(moth_dat) #number of rows or sample observations
m = 100 #number of bootstrap iterations
moth_result = matrix(
  nrow = m,
  ncol = n)

# The outer loop: runs once for each bootstrap iteration.  index variable is i
for(i in 1:m)
{
  # The inner loop: simulates increasing sampling intensity
  # Sampling intensity ranges from 1 site to the complete count of sites (24)
  # index variable is j
  for(j in 1:n)
  {
    # sample the input data row indices, with replacement
    rows_j = sample(n, size = j, replace=TRUE)

    # Creates a new data matrix from the resampled rows.
    t1 = moth_dat[rows_j, ]

    # Calculates the column sums of the new data matrix.
    t2 = apply(t1, 2, sum)

    # Counts the number of columns in which any moths were observed
    moth_result[i, j] = sum(t2 > 0)
  }
}

head(moth_result)

```

```{r include=FALSE, eval=FALSE}

# Packaging your code into a function
# First draft

rarefaction_sampler = function(input_dat, n_iterations)
{
  n = nrow(moth_dat) #number of rows or sample observations
  m = 100 #number of bootstrap iterations

  moth_result = matrix(
    nrow = m,
    ncol = n)

  # The outer loop: runs once for each bootstrap iteration.  index variable is i
  for(i in 1:m)
  {
    # The inner loop: simulates increasing sampling intensity
    # Sampling intensity ranges from 1 site to the complete count of sites (24)
    # index variable is j
    for(j in 1:n)
    {

      # sample the input data row indices, with replacement
      rows_j = sample(n, size = j, replace=TRUE)

      # Creates a new data matrix
      t1 = moth_dat[rows_j, ]

      # Calculates the column sums
      t2 = apply(t1, 2, sum)

      # Counts the number of columns in which any moths were observed
      moth_result[i, j] = sum(t2 > 0)
    }
  }

  return(moth_result)
}

rarefact = rarefaction_sampler(moth_dat, 100)
head(rarefact)




# Second draft

rarefaction_sampler = function(input_dat, n_iterations)
{
  n_input_rows = nrow(input_dat)

  results_out = matrix(
    nrow = n_iterations,
    ncol = n_input_rows)

  # The outer loop: runs once for each bootstrap iteration.  index variable is i
  for(i in 1:n_iterations)
  {
    # The inner loop: simulates increasing sampling intensity
    # Sampling intensity ranges from 1 site to the complete count of
    # sites in the input data (n)
    # index variable is j
    for(j in 1:n)
    {
      # sample the input data row indices, with replacement
      rows_j = sample(n, size = j, replace=TRUE)

      # Creates a new data matrix
      t1 = input_dat[rows_j, ]

      # Calculates the column sums
      t2 = apply(t1, 2, sum)

      # Counts the number of columns in which any moths were observed
      results_out[i, j] = sum(t2 > 0)
    }
  }
  return(results_out)
}

rarefact = rarefaction_sampler(moth_dat, 100)
head(rarefact)




# Check in a fresh environment


# This clears the current R session's environment
rm(list = ls())

# Re-read my data:
moths = read.csv(here("data", "moths.csv"))
moth_dat = moths[,-1]

rarefaction_sampler = function(input_dat, n_iterations)
{
  n_input_rows = nrow(input_dat)

  results_out = matrix(
    nrow = n_iterations,
    ncol = n_input_rows)

  # The outer loop: runs once for each bootstrap iteration.  index variable is i
  for(i in 1:n_iterations)
  {
    # The inner loop: simulates increasing sampling intensity
    # Sampling intensity ranges from 1 site to the complete count of
    # sites in the input data (n)
    for(j in 1:n)
    {
      # sample the input data row indices, with replacement
      rows_j = sample(n, size = j, replace=TRUE)

      # Creates a new data matrix
      t1 = input_dat[rows_j, ]

      # Calculates the column sums
      t2 = apply(t1, 2, sum)

      # Counts the number of columns in which any moths were observed
      results_out[i, j] = sum(t2 > 0)
    }
  }
  return(results_out)
}

rarefact = rarefaction_sampler(moth_dat, 100)
head(rarefact)
```


```{r include=FALSE, eval= FALSE}
# Debugging template

# This clears the current R session's environment
rm(list = ls())

# Re-read my data:
moths = read.csv(here("data", "moths.csv"))

rarefaction_sampler = function(input_dat, n_iterations)
{
  n_input_rows = nrow(input_dat)

  results_out = matrix(
    nrow = n_iterations,
    ncol = n_input_rows)

  # The outer loop: runs once for each bootstrap iteration.  index variable is i
  for(i in 1:n_iterations)
  {
    # The inner loop: simulates increasing sampling intensity
    # Sampling intensity ranges from 1 site to the complete count of
    # sites in the input data (n)
    for(j in 1:n_input_rows)
    {
      # sample the input data row indices, with replacement
      rows_j = sample(n_input_rows, size = j, replace=TRUE)

      # Creates a new data matrix
      t1 = input_dat[rows_j, ]

      # Calculates the column sums
      t2 = apply(t1, 2, sum)

      # Counts the number of columns in which any moths were observed
      results_out[i, j] = sum(t2 > 0)
    }
  }
  return(results_out)
}

rarefact = rarefaction_sampler(moths[,-1], 100)
head(rarefact)


### BUILDING A RAREFACTION CURVE

# Re-read my data:
moths = read.csv(here("data", "moths.csv"))
rarefact = rarefaction_sampler(moths[,-1], 10000)

rare_mean = apply(rarefact, 2, mean)
rare_quant = apply(rarefact, 2, quantile, probs=c(0.025, 0.975))
rare = t(rbind(rare_mean, rare_quant))

# Plotting the curve
matplot(
  rare,
  type='l',
  xlab='Number of sampling plots',
  ylab='Species richness',
  main='Rarefaction Curve')

legend(
  'bottomright',
  legend=c('mean','2.5%','97.5%'),
  lty=c(1,2,3),col=c(1,2,3), inset=c(.1,.1))
```

################################################################################


ASSIGNMENT

```{r}
require(palmerpenguins)
```


1. 123
```{r}
Gentoo = subset(penguins, penguins$species == "Gentoo")
Gentoo_bill = subset(Gentoo, Gentoo$bill_length_mm != "NA")
length(Gentoo_bill$bill_length_mm)
```

2. 3.081857
```{r}
Gentoo_bill_sd = sd(Gentoo_bill$bill_length_mm)
```

3.  1.979439 and -1.979439
```{r}
# Gentoo mean bill length
Gentoo_mean_bill = mean(Gentoo_bill$bill_length_mm)
Gentoo_mean_bill
# Critical-T values
Gentoo_bill_crit_ts = c(qt(1 - 0.05/2, 123), qt(0.05/2, 123))
Gentoo_bill_crit_ts

#Just the positive critical-T
Gentoo_bill_crit_t = qt(1 - 0.05/2, 123)
Gentoo_bill_crit_t
```
4. 0.2778817
```{r}
Gentoo_bill_sse = Gentoo_bill_sd/sqrt(123)
Gentoo_bill_sse
```

5. 47.50488 +/- 0.5500498 
```{r}
# CI radius
Gentoo_bill_CI_radius = Gentoo_bill_crit_t * Gentoo_bill_sse
Gentoo_bill_CI_radius

# CIs: abs() used because the CI radius is negitive
Gentoo_bill_lowCI = Gentoo_mean_bill - Gentoo_bill_CI_radius
Gentoo_bill_upCI = Gentoo_mean_bill + Gentoo_bill_CI_radius

Gentoo_bill_lowCI
Gentoo_bill_upCI
Gentoo_bill_CIs = c(Gentoo_bill_lowCI, Gentoo_bill_upCI)
Gentoo_bill_CIs


### OR 


Gentoo_bill_CI = c(Gentoo_mean_bill + Gentoo_bill_CI_radius, 
  Gentoo_mean_bill - Gentoo_bill_CI_radius)
Gentoo_bill_CI
```

6. 47.50488 +/- 0.544 (The mean -/+ 2.5% or 97.5% quantile)

7. 
```{r}
require(boot)

boot_mean = function(x, i)
{
  return(mean(x[i], na.rm = TRUE))
}


Gentoo_boot = 
  boot(
    data = Gentoo_bill$bill_length_mm,
    statistic = boot_mean,
    R = 10000)
print(Gentoo_boot)
```

8. 
```{r}
quantile(
  Gentoo_boot$t,
  c(0.025, 0.975))
```


9.
```{r}
rarefaction_sampler = function(input_dat, n_iterations)
{
  n_input_rows = nrow(input_dat)

  results_out = matrix(
    nrow = n_iterations,
    ncol = n_input_rows)

  # The outer loop: runs once for each bootstrap iteration.  index variable is i
  for(i in 1:n_iterations)
  {
    # The inner loop: simulates increasing sampling intensity
    # Sampling intensity ranges from 1 site to the complete count of
    # sites in the input data (n)
    for(j in 1:n_input_rows)
    {
      # sample the input data row indices, with replacement
      rows_j = sample(n_input_rows, size = j, replace=TRUE)

      # Creates a new data matrix
      t1 = input_dat[rows_j, ]

      # Calculates the column sums
      t2 = apply(t1, 2, sum)

      # Counts the number of columns in which any moths were observed
      results_out[i, j] = sum(t2 > 0)
    }
  }
  return(results_out)
}
```

10. I had a difficult time understanding what each of the components in the function were doing. This in turn made it difficult to understand how to correct the error. I still do not think that I fully understand what exactly is going on. 

11. 
```{r}
moths = read.csv(here("data", "moths.csv"))
rarefact = rarefaction_sampler(moths[,-1], 10000)


rare_mean = apply(rarefact, 2, mean)
rare_quant = apply(rarefact, 2, quantile, probs=c(0.025, 0.975))
rare = t(rbind(rare_mean, rare_quant))
```

12. 
```{r}
# Plotting the curve
matplot(
  rare,
  type='l',
  xlab='Number of Sampling Plots',
  ylab='Species Richness',
  main= "JT's Rarefaction Curve",
  lty = c(1, 2, 2),
  lwd = c(2,4,4),
  col = c(1, 4, 6)
  )

legend(
  'bottom',
  legend=c('Average Number of Species Present','2.5% Quantile','97.5% Quantile'),
  lty=c(1,2,2),col=c(1,4,6), lwd = c(2,4,4), inset=c(.1,.1))
```
 
13. We should visit at least 30 sites to see all of the moth species.This is where species richness is highest and where the 2.5 quantile, 97.5 quantile, and the mean converge. Additionally, we can apply the central limit to our data and assume we are approaching the true number of species present at our sites.


