---
title: "Lab 6_Introduction To Inferance"
author: "JT Larkin (Partners: Sonja Glasser, Ethan Rutledge)"
date: "10/20/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include=FALSE}
##Practice


require(palmerpenguins)

sse_mean = function(x)
    { na.x<-is.na(x)
          x2<- x[!na.x]
                 sd.x2=sd(x2)
                  n=length(x2)
       sse = sd.x2/(sqrt(n))
       return(sse)
}


sse_mean(penguins$bill_depth_mm)







boxplot(flipper_length_mm ~ species, data = penguins)

dat_pen = subset(penguins, species != "Gentoo")
boxplot(flipper_length_mm ~ species, data = dat_pen)




dat_pen = droplevels(subset(penguins, species != "Gentoo"))
{
  par(mfrow = c(1, 2))
  boxplot(flipper_length_mm ~ species, data = penguins)
  boxplot(flipper_length_mm ~ species, data = dat_pen)
}



# for reproducibility
set.seed(123)

flipper_shuffled = sample(penguins$flipper_length_mm, replace = TRUE)
par(mfrow = c(1, 2))
boxplot(flipper_length_mm ~ species, data = penguins)
boxplot(flipper_shuffled ~ penguins$species, xlab = "species")

##Classical t-test: Adelie and Chinstrap penguins

t.test(dat_pen$flipper_length_mm ~ dat_pen$species)


# Reset the random number generator state for reproduceablility
set.seed(1)
flipper_shuffled = sample(dat_pen$flipper_length_mm)


boxplot(flipper_shuffled ~ dat_pen$species)


#T-test on resampled data
t_test_1 = t.test(flipper_shuffled ~ dat_pen$species)
t_test_1


##Differance of means

t_test = t.test(dat_pen$flipper_length_mm ~ dat_pen$species)
t_test

t_test$estimate

diff_observed = round(diff(t_test$estimate), digits = 3)
print(diff_observed, digits = 3)



#Using aggregate

agg_means = aggregate(
  flipper_length_mm ~ species, 
  data = dat_pen, 
  FUN = "mean", 
  na.rm = TRUE)
diff_observed = diff(agg_means[, 2])

agg_means
diff_observed



##Sample sizes
table(dat_pen$species)

n_1 = 68
n_2 = 152

dat_1 = sample(dat_pen$flipper_length_mm, n_1, replace = TRUE)
dat_2 = sample(dat_pen$flipper_length_mm, n_2, replace = TRUE)

diff_simulated = 
  mean(dat_1, na.rm = TRUE) - mean(dat_2, na.rm = TRUE)

print(c(observed = diff_observed, simulated = diff_simulated))


##Simulation Function
x = dat_pen$flipper_length_mm
n_1 = 68
n_2 = 152

dat_1 = sample(x, n_1, replace = TRUE)
dat_2 = sample(x, n_2, replace = TRUE)

diff_simulated = 
  mean(dat_1, na.rm = TRUE) - mean(dat_2, na.rm = TRUE)



two_group_resample = function(x, n_1, n_2)
  {
  dat_1 = sample(x, n_1, replace = TRUE)
  dat_2 = sample(x, n_2, replace = TRUE)

  diff_simulated = 
  mean(dat_1, na.rm = TRUE) - mean(dat_2, na.rm = TRUE)
 
  return(diff_simulated) 
  }


set.seed(54321)
two_group_resample(dat_pen$flipper_length_mm, 68, 152)



#Resampling experiment

n = 200
mean_differences = c()
for (i in 1:n)
{
  mean_differences = c(
    mean_differences,
    two_group_resample(dat_pen$flipper_length_mm, 68, 152)
  )
}
hist(mean_differences)

sum(abs(mean_differences) >= diff_observed)



#Retrieving named elements

t_test = t.test(flipper_shuffled ~ dat_pen$species)
str(t_test)
t_test$estimate
```


################################################################################

ASSIGNMENT


1. 
```{r}
rm(list = ls())
sse_mean = function(x)
    { na.x<-is.na(x)
          x2<- x[!na.x]
                 sd.x2=sd(x2)
                  n=length(x2)
       sse = sd.x2/(sqrt(n))
       return(sse)
}

sse_mean(penguins$body_mass_g)
sse_mean(mtcars$mpg)
```

2.
```{r}
two_group_resample = function(x, n_1, n_2)
  {
  dat_1 = sample(x, n_1, replace = TRUE)
  dat_2 = sample(x, n_2, replace = TRUE)

  diff_simulated = 
  mean(dat_1, na.rm = TRUE) - mean(dat_2, na.rm = TRUE)
 
  return(diff_simulated) 
  }
```

3. Our function is preforming a Monte Carlo resampling method which is simulating a null hypothesis (no difference between penguin mean flipper lengths). It shuffles each column separately so that the data is completely random.

4.
```{r}
dat_pen2 = droplevels(subset(penguins, species != "Gentoo"))


n = 2000
mean_differences = c()
for (i in 1:n)
{
  mean_differences = c(
    mean_differences,
    two_group_resample(dat_pen2$flipper_length_mm, 68, 152)
  )
}
hist(mean_differences,
     xlab = "Mean differences of simulated penguin mass data",
     ylab = "Frequency",
     main = "Histogram of mean differences \nfrom simulated data")
```

5. 0 of the resampled differences of means were greater than or equal to 5.8. Our p-value was very, very small meaning our finding was very significant. 

```{r}
t.test2 = t.test(dat_pen2$flipper_length_mm~dat_pen2$species)
t.test2
sum(abs(mean_differences) >= 5.8)
```

6. It will be a lot. It is not possible to give an exact number. It may occur at 10 million simulations or it may not occur until 25 million or 50 million simulations. Regardless, it will be a very long time because it is a very, very rare occurrence. 

7.

```{r}
boxplot(dat_pen2$body_mass_g~dat_pen2$species, 
        xlab = "Species",
        ylab = "Mass in grams (g)",
        main = "Boxplots of 2 Penguin Species Mass's")

```

8. 
Group Means
- Adellie: 3700.662 g
- Chinstrap: 3733.088 g

Difference Between Means
- 32.42598

```{r}
agg_means = aggregate(
  body_mass_g ~ species, 
  data = dat_pen2, 
  FUN = "mean", 
  na.rm = TRUE)
diff_crit = diff(agg_means[, 2])

agg_means
diff_crit
```

9. When running a t-test, we have 2 hypotheses: Null and the Alternative. The null was that there is no difference in mean mass between Adelie and Chinstrap penguins. The alternative is that there is a difference in mean mass between the 2 species. Since this is a 2-tailed t-test, we do not specify how they differ (e.g. Chinstrap penguins are heavier than Adelie), but rather just say that there is a difference. For a p-value less than 0.05 we accept the alternative, but for p-values greater than 0.05 we accept the null. We have a p-value of 0.5879, thus we accept the null that there is no difference in mean mass between the 2 species. About 60 percent of the time we will observe a difference of at least or greater than 32.42 grams.
```{r}
t.test(dat_pen2$body_mass_g ~ dat_pen2$species)

```

10. 613
```{r}
n = 1000
mean_differences2 = c()
for (i in 1:n)
{
  mean_differences2 = c(
    mean_differences2,
    two_group_resample(dat_pen2$body_mass_g, 68, 152)
  )
}



t.test2 = t.test(dat_pen2$body_mass_g~dat_pen2$species)
sum(abs(mean_differences2) >= diff_crit)
```

11.
```{r}
hist(mean_differences2,
     xlab = "Mean differences of simulated penguin mass data",
     ylab = "Frequency",
     main = "Histogram of mean differences \nfrom simulated data")
```

