---
title: "Modeling Data 2"
author: "JT Larkin"
date: "11/10/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
require(here)
catrate = read.csv(here("data", "catrate.csv"))
head(catrate)
```

```{r include=FALSE}
# Binomial Test for Proportions

## Reproductive Success and Failure

n_success = sum(catrate$success)
n_years = sum(catrate$years)
binom.test(n_success, n_years)

## Reproductive Catastrophe and Late Filling

binom.test(n_success, n_years, p = 5/7) 
```
```{r include=FALSE}
# Two Sample Tests

## F-distribution Example: Vegetation Data

veg = read.csv(here("data", "vegdata.csv"))
head(veg)

boxplot(pine ~ treatment, data = veg)

## Variance Test

var.test(
  pine ~ treatment,
  data = veg,
  subset = treatment %in% c('control','clipped'))

## F-tests Assumes Normality

shapiro.test(veg$pine[veg$treatment=="control"])

shapiro.test(veg$pine[veg$treatment=="clipped"])

```

```{r include=FALSE}
# Non-parametric Variance Test

fligner.test(
  pine ~ treatment,
  data = veg,
  subset = treatment %in% c('control','clipped'))

bartlett.test(pine ~ treatment, data=veg)


fligner.test(pine ~ treatment, data = veg)


```

```{r include=FALSE}
# Comparing 2 sample means

## T-test

t.test(pine~treatment,data=veg,subset=treatment %in% c('control','clipped'), conf.int=TRUE)


## Wilcox test

wilcox.test(pine~treatment,data=veg,subset=treatment %in% c('control','clipped'), conf.int=TRUE)


```

```{r include=FALSE}
# Tests for paired samples

control = veg$pine[veg$treatment=='control']
clipped = veg$pine[veg$treatment=='clipped']


t.test(control, clipped, paired=TRUE)
wilcox.test(control, clipped, paired=TRUE)

```


```{r include= FALSE}
# Correlation

disp = read.csv(here("data", "dispersal.csv"))
disp

plot(disp$disp.rate.ftb, disp$disp.rate.eb)

cor.test(
  disp$disp.rate.ftb,
  disp$disp.rate.eb,
  use='complete.obs')

cor.test(
  disp$disp.rate.ftb,
  disp$disp.rate.eb,
  use='complete.obs',
  method='spearman')



# Comparing two distributions

plot(
  ecdf(disp$disp.rate.ftb),
  verticals=TRUE)



plot(
  ecdf(disp$disp.rate.ftb),
  verticals=TRUE)
plot(
  ecdf(disp$disp.rate.eb),
  verticals=TRUE,
  lty=3,
  add=TRUE)



ks.test(disp$disp.rate.ftb,disp$disp.rate.eb)

prop.test(c(4,16),c(40,250))


```

```{r include=FALSE}
# Contingency: Chi-square test

owls = matrix(c(16, 9, 4, 11), nrow=2)
rownames(owls) = c("present", "absent")
colnames(owls) = c("old", "young")
chisq.test(owls)


## Fishers exact test

fisher.test(owls)



## Bird hab data

birds   = read.csv(here("data", "bird.sta.csv"))
hab     = read.csv(here("data", "hab.sta.csv"))
birdhab = merge(birds, hab, by=c("basin", "sub", "sta"))

# Create a contingency table for edge/interior and brown creeper presence/absence
table(birdhab$s.edge, birdhab$BRCR > 0)

# set the presence to be in the first column
br_creeper_table = table(birdhab$s.edge, birdhab$BRCR > 0)[, 2:1]
chisq.test(br_creeper_table)

```


################################################################################

ASSIGNMENT

1. There is no relationship between brown creeper presence/absence and edge/interior habitats.

2. I think that brown creepers show a habitat preference. Our p-value from the chi squared test was very significant. Thus, the null hypothesis (no relationship) is rejected.  

```{r}
chisq.test(br_creeper_table)
```

3. 

```{r}
require(palmerpenguins)
fit_species = 
  lm(
    formula = body_mass_g ~ species,
    data = penguins)
```

4. 
```{r}
fit_sex= 
  lm(
    formula = body_mass_g ~ sex,
    data = penguins)
```

5. 
```{r}
fit_both = 
  lm(
    formula = body_mass_g ~ species * sex,
    data = penguins)
```

6. 

```{r}
levels(penguins$species)

boxplot(body_mass_g ~ species,
    data = penguins, 
  main = "Plot of Body Mass By Penguin Species", 
   xlab = "",
  ylab = "Body Mass (g)",
  names = c("Adelie", "Chinstrap", "Gentoo"))
```

7. 
```{r}
levels(penguins$sex)

boxplot(body_mass_g ~ sex,
    data = penguins, 
  main = "Plot of Body Mass By Penguin Sex",
   xlab = "",
  ylab = "Body Mass (g)",
  names = c("Female", "Male"))
```

8. 
```{r}
boxplot(body_mass_g ~ species * sex,
    data = penguins, 
  main = "Plot of Body Mass By Penguin Species and Sex",
  xlab = "",
  ylab = "Body Mass (g)",
  las = 2,
  names = c("Adelie \nFemale", "Adelie \nMale", "Chinstrap \nFemale",  
            "Chinstrap \nMale","Gentoo \nFemale",  "Gentoo \nMale"))
```


9. I would say that the fit_species may have a problem fulfilling that homogeneity assumption.The variability between the species body mass's is pretty clear in the boxplot we created in question 6.

10. There is no difference in variances between groups. 

11. p-value = 0.0501

```{r}
bartlett.test(body_mass_g ~ species, data = penguins)
```

12. p-value = 0.0319
```{r}
bartlett.test(body_mass_g ~ sex, data = penguins)
```


13. p-value = 0.1741
```{r}
dat_groups = aggregate(
  body_mass_g ~ species * sex,
  data = penguins,
  FUN = c)
str(dat_groups)


bartlett.test(dat_groups$body_mass_g)

```

14. Since our p-value is not significant for the two-way model, we will assert the null hypothesis that there is no difference in variance between groups.Thus, we will not have an issue with heterogeneity in the two-way model. The p-value from the Bartlett test on the sex model is significant. Thus, we will assert the alternative hypothesis that states there is a difference in variance between sexes and we will have an issue with heterogeneity. Finally, the p-value from the Bartlett test for the species model had a p-value that is right on the border of significance (p-value = 0.05001). Thus, drawing a conclusion on whether this model will have issues with heterogeneity is difficult.








