---
title: 'Lab 1: R Fundamentals 1'
author: "JT Larkin"
subtitle: "Partners: Ethan Rutledge"
date: "9/9/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
```{r}
c(1, 2, 3)

"c(1, 2, 3)"

```

1. Explain why the outputs of the two lines are different?

  *	It is seeing c(1,2,3) as numeric, where as "c(1,2,3)" is seen as text.


```{r}
c_1 = c(1, 2, 3)
c_2 = "c(1, 2, 3)"
```

2. Is c_1 a variable, or a function?

  *	Variable. We are using the combine function (c()) to give the variable a value.

3. Is c_2 a variable, or a function?

  *	c_2 is also a variable. We are assigning the character value "c(1, 2, 3)" to it.

4. If c_1 and c_2 have different values, why?

  * Yes, they have different values. This is because c_1 is numeric and c_2 is a character.


```{r}
my_vec=c(1:3)
mat_1 = matrix(my_vec)
mat_1

mat_1[[3]]
```


5. What are the dimensions of the matrix (i.e. how many rows and columns)?

  * It is a matrix with one column and 3 rows


6. Write R code to retrieve the element of mat_1 that has a value of 3.

  * mat_1[[3]]


```{r}
mat_2= matrix(my_vec, nrow= 2, ncol=3)
mat_2

mat_3= matrix(my_vec, nrow= 3, ncol=2)
mat_3

mat_4= matrix(my_vec, nrow=2, ncol=4)
mat_4
```


7. Paste the code you used to create mat_2.

  * mat_2= matrix(my_vec, nrow= 2, ncol=3)

8. Paste the code you used to create mat_3.

  * mat_3= matrix(my_vec, nrow= 3, ncol=2)

9. Did R use rows or columns to recycle the values in my_vec?

  * Columns

10. Create a matrix, mat_4, with a number of elements that is not a multiple of 3 and paste the code into the editor.

  * mat_4= matrix(my_vec, nrow=2, ncol=4)

11. How did R handle the recycling of values of my_vec in mat_4?

  * It created the matrix, but returned a warning code: "data length [3] is not a sub-multiple or multiple of the number of rows [2]".It only used the numbers 1 and 2 on the 3rd recycle.


```{r}
my_list_1= list(two=5.2, one="five point two",three=0:5)
my_list_1


my_list_1[[1]]
my_list_1[[as.numeric("1")]]
my_list_1[["1"]]
my_list_1[["one"]]
my_list_1$one
my_list_1$"one"
#my_list_1$1
my_list_1$"1"

```


12. For each of the 8 lines, answer the following: A. Did the line return a 1: value, 2: error, or 3: NULL B. If it did not return an error, what type of subsetting operation was used? C. If it did not 
return an error, explain how R chose which element to retrieve.

  * 1- a) value; b) [[]]; c) It "grabbed" the first element in the list. 
  
  * 2- a) value; b) [[]]; c) It "grabbed" the first element in the list and displayed it as the numeric value.
  
  * 3- a) NULL; b) [[]]; c) It is looking for something labeled "1" (text), which does not exist
  
  * 4- a) value; b) [[]]; c) It "grabbed" the element that we labeled "one" and returned its value
  
  * 5- a) value; b) $; c) It "grabbed" the element labeled "one"

  * 6- a) value; b) $; c) This also "grabbed" the value labeled "one"

  * 7- a) Error; b) $; c) N/A

  * 8- a) Null; b) $; c) It is looking for something labeled "1" (text), which does not exist



13. Identify which lines produced output "five point two" and explain why.

  * 4,5, and 6. All three of these lines are searching for the element in our vector we named "one" which was "five point two". 

14. Identify which lines produced NULL output and explain why.

  * 4 and 8. R was looking for something labeled "1" (text), which does not exist in the vector my_list_1
