---
title: 'Lab 11: Simulation and Power Analysis'
author: "JT Larkin (Partners: Alex Fink and Sonja Glasser)"
date: "11/30/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
require(here)
bird = read.csv(here("data", "bird.sub.csv"))
hab = read.csv(here("data", "hab.sub.csv"))
bird
hab


birdhab = merge(bird, hab, by = c("basin" , "sub"))
birdhab

dim(birdhab)
```


```{r include=FALSE}
plot(data = birdhab, BRCR ~ ls)
fit_1 = lm(data = birdhab, BRCR ~ ls)
abline(fit_1)

summary(fit_1)
```

```{r include=FALSE}
#Deterministic Model: Linear Function

linear = function(x, y_int, slope)
{
 y = slope * x + y_int
 return(y) 
}


#self check

linear(x = 1, y_int = 1, slope = 1)
linear(x = 3:5, y_int = 1, slope = 1)
linear(x = 3:5, y_int = -1, slope = 1)
linear(x = 3:5, y_int = -1, slope = 0.01)


#Stochastic Model
#rnorm(n, mean=0, st_dev)
```

```{r include=FALSE}
#Simulation Function


linear_simulator = function(x, y_int, slope, st_dev)
{
 n = length(x)
 (y = slope * x + y_int) + (rnorm(n, mean = 0, st_dev))
}



#Self test 1
n = 200

par(mfrow = c(2, 2))
for (i in 1:4)
{
  x = runif(n = n)
  plot(
    x,
    linear_simulator(x, y_int = 1, slope = 4.5, st_dev = 0.1),
    main = "", xlab = "x", ylab = "y",
    pch = 16, col = rgb(0, 0.2, 0, 0.2))
}


#self test 2
png(filename = here("images", "Lab", "Lab11_plot.png"), width = 1500, height = 1600, res = 180)

n = 400

par(mfrow = c(2, 2))
for (i in 1:4)
{
  x = runif(n = n)
  plot(
    x, linear_simulator(x, y_int = 10, slope = -6.5, st_dev = 1.1),
    main = "", xlab = "x", ylab = "y",
    pch = 16, col = rgb(0, 0.2, 0, 0.2))
}

dev.off()
```


```{r include=FALSE}
fit_1_coefs = coefficients(fit_1)
str(fit_1_coefs)



fit_1_summary = summary(fit_1)
str(fit_1_summary)
fit_1_summary$sigma
fit_1_summary


int_obs =  0.0991039 
slope_obs = 0.0058405
sd_obs = 0.1412668
```

```{r include=FALSE}
plot(
  x = birdhab$ls, 
  y = linear_simulator(
    x = birdhab$ls,
    y_int = int_obs,
    slope = slope_obs,
    st_dev = sd_obs
  ),
  main = "Simulated Data",
  xlab = "late-successional forest",
  ylab = "Brown Creeper Abundance")
```


```{r include=FALSE}
plot(
  birdhab$ls, birdhab$BRCR, 
  xlab = "late-successional forest extent",
  ylab = "Brown Creeper abundance",
  pch = 19)

points(
  x = birdhab$ls, 
  y = linear_simulator(
    x = birdhab$ls,
    y_int = int_obs,
    slope = slope_obs,
    st_dev = sd_obs
  ),
  col = adjustcolor("red", alpha = 0.3),
  pch = 16)

legend(
  "topleft",
  legend = c("data", "simulation"),
  pch = 16,
  col = c(1, adjustcolor("red", alpha = 0.3)))
```

```{r include=FALSE}
#Single Simulation
y_sim = linear_simulator(
  x = birdhab$ls,
  y_int = int_obs,
  slope = slope_obs,
  st_dev = sd_obs
)

fit_sim = lm(y_sim ~ birdhab$ls)
summary(fit_sim)




sum_1 = summary(fit_sim)
sum_1$coefficients
```


```{r include=FALSE}
n_sims = 1000
p_vals = numeric(n_sims)
alpha = 0.05
for(i in 1:n_sims)
{
  y_sim = linear_simulator(
    x = birdhab$ls,
    y_int = int_obs,
    slope = slope_obs,
    st_dev = sd_obs
  )
  fit_sim = lm(y_sim ~ birdhab$ls)
  
  p_vals[i] = summary(fit_sim)$coefficients[2, 'Pr(>|t|)']
}
sum(p_vals < alpha) / n_sims
```


```{r include=FALSE}
linear_sim_fit = function(x, y, slope, y_int, st_dev)
{
  y_sim = linear_simulator(
    x = x,
    y_int = y_int,
    slope = slope,
    st_dev = st_dev
  )
  fit_sim = lm(y_sim ~ x)
  return(fit_sim)
}
```


```{r include=FALSE}
alpha = 0.05
n_sims = 1000
p_vals = numeric(n_sims)

n_effect_sizes = 20
effect_sizes_1 = seq(-.01, .01, length.out = n_effect_sizes)

effect_size_powers = numeric(n_effect_sizes)

for(j in 1:n_effect_sizes)
{
  for(i in 1:n_sims)
  {
    fit_sim = linear_sim_fit(
      x = birdhab$ls,
      y_int = int_obs,
      slope = effect_sizes_1[j],
      st_dev = sd_obs
    )
    
    p_vals[i] = summary(fit_sim)$coefficients[2, 'Pr(>|t|)']
  }
  effect_size_powers[j] = sum(p_vals < alpha) / n_sims
}

sim_effect_size = 
  data.frame(
    power       = effect_size_powers,
    effect_size = effect_sizes_1)
```



```{r include=FALSE}
plot(
  power ~ effect_size, data = sim_effect_size,
  type = 'l', xlab = 'Effect size', ylab = 'Power')
abline(v = slope_obs, lty = 2, col = 'red')
```




```{r include=FALSE}
alpha = 0.05
n_sims = 1000
p_vals = numeric(n_sims)

sample_sizes = seq(5, 100)
sample_size_powers = numeric(length(sample_sizes))

for(j in 1:length(sample_sizes))
{
  x_vals = seq(0, 100, length.out = sample_sizes[j])
  
  for(i in 1:n_sims)
  {
    fit_sim = linear_sim_fit(
      x = x_vals,
      y_int = int_obs,
      slope = slope_obs,
      st_dev = sd_obs
    )
    p_vals[i] = summary(fit_sim)$coefficients[2, 'Pr(>|t|)']
  }
  sample_size_powers[j] = sum(p_vals < alpha) / n_sims
}

sim_sample_size = 
  data.frame(
    power = sample_size_powers,
    sample_size = sample_sizes)


plot(
  power ~ sample_size, data = sim_sample_size,
  type = 'l', xlab = 'Sample size', ylab = 'Power')
abline(v = nrow(birdhab), lty = 2, col = 'red')
```



```{r include=FALSE}
alpha = 0.01
n_sims = 50

p_vals = numeric(n_sims)

n_effect_sizes = 20
effect_sizes = seq(-.01, .01, length.out = n_effect_sizes)

sample_sizes = seq(10, 50)

sim_output_2 = matrix(nrow = length(effect_sizes), ncol = length(sample_sizes))

for(k in 1:length(effect_sizes))
{
  effect_size = effect_sizes[k]
  for(j in 1:length(sample_sizes))
  {
    x_vals = seq(0, 100, length.out = sample_sizes[j])
    
    for(i in 1:n_sims)
    {
      fit_sim = linear_sim_fit(
        x = x_vals,
        y_int = int_obs,
        slope = effect_size,
        st_dev = sd_obs
      )
      p_vals[i] = summary(fit_sim)$coefficients[2, 'Pr(>|t|)']
    }
    sim_output_2[k, j] = sum(p_vals < alpha) / n_sims
  }
  print(paste0("computing effect size ", k," of ", length(effect_sizes)))
}

sim_n_effect_size = 
  list(
    power = sim_output_2,
    effect_size = effect_sizes,
    sample_size = sample_sizes
  )

```


```{r include=FALSE}
image(
  sim_n_effect_size$power,
  xlab = "Effect size",
  ylab = "Sample Size",
  axes = FALSE)

# add x-axis labels
axis(
  1, 
  at = c(0, 0.5, 1), 
  labels = c(-.01, 0.0, .01))

# add y=axis labels
axis(
  2, 
  at = c(0, 1), 
  labels = c(sample_sizes[1], tail(sample_sizes, 1)))
```



```{r include=FALSE}
contour(
  x = sim_n_effect_size$effect_size,
  y = sim_n_effect_size$sample_size,
  z = sim_n_effect_size$power,
  xlab = "effect size",
  ylab = "sample size",
  main = "Contour Plot of Statistical Power",
  levels = seq(0, 1, length.out = 9),
  drawlabels = TRUE,
  # method = "simple")
  method = "edge")
```



```{r include=FALSE}
persp(
  x = sim_n_effect_size$effect_size,
  y = sim_n_effect_size$sample_size,
  z = sim_n_effect_size$power,
  xlab = "beta", ylab = "n", zlab = "power",
  col = 'lightblue',
  theta = 30, phi = 30, expand = .75,
  ticktype = 'detailed')
```
```{r include=FALSE}
require(rgl)
persp3d( x = sim_n_effect_size$effect_size,
  y = sim_n_effect_size$sample_size,
  z = sim_n_effect_size$power,
  xlab = "beta", ylab = "n", zlab = "power",
  col = 'lightblue',
  theta = 30, phi = 30, expand = .75,
  ticktype = 'detailed')
```




```{r include=FALSE, eval=FALSE}
require(htmlwidgets)
saveWidget(
  rglwidget(),
  file = here(
    "docs",
    "n_effect_size_power_sim_plot.html"),
  selfcontained = TRUE
)
```


```{r include=FALSE}
save(
  sim_n_effect_size,
  file = here::here("data", "lab_11_n_effect_sizes.Rdata"))

load(file = here::here("data", "lab_11_n_effect_sizes.Rdata"))
```


```{r include=FALSE}
#Population Dispersion Analysis

require(here)

alpha = 0.05
n_sims = 100
p_vals = numeric(n_sims)

# What was the observed standard deviation?
sd_obs

# specify the number of different standard deviation values to simulate:
n_sds = 20
pop_sds = seq(from = 0.01, to = 1.5, length.out = n_sds)

pop_sd_power = numeric(n_sds)

for(j in 1:length(pop_sds))
{
  pop_sd_j = pop_sds[j]
  for(i in 1:n_sims)
  {
    fit_sim = linear_sim_fit(
      x = birdhab$ls,
      y_int = int_obs,
      slope = slope_obs,
      st_dev = pop_sd_j
    )
    p_vals[i] = summary(fit_sim)$coefficients[2, 'Pr(>|t|)']
  }
  pop_sd_power[j] = sum(p_vals < alpha) / n_sims
}

sim_output_dispersion = data.frame(
  sd = pop_sds,
  power = pop_sd_power)

# You should save your simulation results so you don't have to run it every time.
save(
  sim_output_dispersion, 
  file = here("data", "lab_ll_dat_dispersion_sim.RData"))

# Line plot of standard deviation (x-axis) and statistical power (y-axis)
plot(power ~ sd, data = sim_output_dispersion,
  type = 'l', xlab = 'Standard Deviation', ylab = 'Power',
  main =  "Population Dispersion Analysis Line Plot")


# Add a dotted vertical red line at the observed population standard deviation value.
abline(v=sd_obs, lty = 2, col = 'red')
```

```{r include=FALSE, eval= FALSE}
#Population Dispersion and Sample Size Analysis

alpha = 0.05

# Start with a small number
n_sims = 100
p_vals = numeric(n_sims)

# What was the observed standard deviation?
sd_obs

# specify the number of different standard deviation values to simulate:
# Start with a small number
n_sds = 20
pop_sds = seq(from = 0.15, to = 1.5, length.out = n_sds)

pop_sd_power = numeric(n_sds)

sample_sizes = seq(5, 100)

sim_output_3 = matrix(0, nrow = n_sds, ncol = length(sample_sizes))

for(k in 1:length(pop_sds))
{
  pop_sd_k = pop_sds[k]
  
  for(j in 1:length(sample_sizes))
  {
    x_vals = seq(0, 100, length.out = sample_sizes[j])
    
    for(i in 1:n_sims)
    {
      fit_sim = linear_sim_fit(
        x = x_vals,
        y_int = int_obs,
        slope = effect_size,
        st_dev =  pop_sd_k
      )
      p_vals[i] =summary(fit_sim)$coefficients[2, 'Pr(>|t|)']
    }
    
    sim_output_3[k, j] = sum(p_vals < alpha) / n_sims
  }
  print(paste0("Testing standard deviation ", k, " of ", n_sds))
}

image(sim_output_3)

sim_3_dat = 
  list(
    power = sim_output_3,
    sample_size = sample_sizes,
    pop_sd = pop_sds)


# You should save your simulation results so you don't have to run it every time.
save(
  sim_3_dat, 
  file = here::here("data", "lab_ll_sim_output_dispersion_n_1000.RData"))

```

################################################################################

Assignment

# Questions 1-2 {.tabset .tabset-pills}
1. 
```{r}
plot(power ~ sd, data = sim_output_dispersion,
  type = 'l', xlab = 'Standard Deviation', ylab = 'Power',
  main =  "Population Dispersion Analysis Line Plot")
abline(v=sd_obs, lty = 2, col = 'red')
```
2. Statistical power decreases as dispersion increases because there is a lower chance that the true effect or value is actually captured.

# Questions 3-4

3. 
```{r}
#Population Dispersion and Sample Size Analysis

alpha = 0.05

# Start with a small number
n_sims = 100
p_vals = numeric(n_sims)

# What was the observed standard deviation?
sd_obs

# specify the number of different standard deviation values to simulate:
# Start with a small number
n_sds = 20
pop_sds = seq(from = 0.15, to = 1.5, length.out = n_sds)

pop_sd_power = numeric(n_sds)

sample_sizes = seq(5, 100)

sim_output_3 = matrix(0, nrow = n_sds, ncol = length(sample_sizes))

for(k in 1:length(pop_sds))
{
  pop_sd_k = pop_sds[k]
  
  for(j in 1:length(sample_sizes))
  {
    x_vals = seq(0, 100, length.out = sample_sizes[j])
    
    for(i in 1:n_sims)
    {
      fit_sim = linear_sim_fit(
        x = x_vals,
        y_int = int_obs,
        slope = effect_size,
        st_dev =  pop_sd_k
      )
      p_vals[i] =summary(fit_sim)$coefficients[2, 'Pr(>|t|)']
    }
    
    sim_output_3[k, j] = sum(p_vals < alpha) / n_sims
  }
  print(paste0("Testing standard deviation ", k, " of ", n_sds))
}

sim_3_dat = 
  list(
    power = sim_output_3,
    sample_size = sample_sizes,
    pop_sd = pop_sds)


# You should save your simulation results so you don't have to run it every time.
save(
  sim_3_dat, 
  file = here::here("data", "lab_ll_sim_output_dispersion_n_1000.RData"))

contour(
  x = sim_3_dat$pop_sd,
  y = sim_3_dat$sample_size,
  z = sim_3_dat$power,
  xlab = "Dispersion",
  ylab = "Sample size",
  main = "Contour Plot of Statistical Power",
  levels = seq(0, 1, length.out = 9),
  drawlabels = TRUE,
  #method = "simple", 
  method = "edge")

```

4. Statistical power increases with lower dispersion and higher sample sizes. Lowest statistical power is found when sample size is low and dispersion is high. This makes sense because as previously stated in question 2, there is a lower chance that the true effect or value is actually captured. 


# Questions 5-6 


5. 
```{r}
require(rgl)
persp3d( x = sim_3_dat$pop_sd,
  y = sim_3_dat$sample_size,
  z = sim_3_dat$power,
  xlab = "beta", ylab = "n", zlab = "power",
  col = 'goldenrod',
  theta = 30, phi = 30, expand = .75,
  ticktype = 'detailed')
```

```{r eval=FALSE}
require(htmlwidgets)
saveWidget(
  rglwidget(),
  file = here(
    "docs",
    "disp_size_power_sim_plot.html"),
  selfcontained = TRUE
)
```


6. You can use the information that is displayed in the 3D plot to determine the number of samples you should collect given your desired statistical power. If I wanted to have a statistical power of at least 0.8, I would know that I would need to collect at least 60 samples with very little dispersion or to be safe around 80 samples which would allow for some dispersion of the data.