---
title: 'Data Camp: Intro to R'
author: 'JT Larkin (Partners: Ethan Rutledge)'
date: "9/8/2021"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
a= c("Jeffery")
a
class(a)
```
1.) What type of data is contained in the variable "a"?

-Character
         
```{r}
b1=c(45.6)
b1
class(b1)
```

2.) What type of data is contained in the variable "b1"?

-Numeric
      
```{r}
b2= c("45.6")
b2
class(b2)
```

3.) What type of data is contained in the variable "b2"?

-Character
        
        
```{r}
#b1 + b2
```

4.) What happens when you add b1 and b2 and why?

-It will return the error message "Error in b1 + b2 : non-numeric argument to binary operator". It returns this because we are trying to add a numeric variable and a character variable, which is not possible.
    
    
```{r}
c1= c(0, 1, 2, 3)
c1
class(c1)
```

5.) Are the variables b1 and c1 of the same type?

-Both are numeric variables. However, b1 is rational number where c1 is a string of integers.

b1= numeric, rational number
c1= numeric, string of integers
      
```{r}
b1 + c1
```

6.) Explain what happens when you add b1 and c1. Consider both the number of elements in each variable and the data types.

-Since both variables are numeric, it adds 45.6 and 0, then it adds 45.6 and 1, next 45.6 and 2, and finally 45.6 and 3. The output looks like this: "[1] 45.6 46.6 47.6 48.6". 
    
    
```{r}
v1= c(-2, -1, 0, 1, 2)
v1
```
    
7.) Show the R code you used to create "v1".

-v1= c(-2, -1, 0, 1, 2)
    
    
```{r}
v2= v1 * 3
v2
```
    
8.) Show the R code you used to create "v2". 

-v2= v1 * 3
    
    
    
```{r}
sum(v2)
```
9.) Show the R code you used to calculate the sum of elements in "v2".

-sum(v2)
      
      
```{r}
vec4= c(1:12)
vec4


mat_1= matrix(vec4, nrow= 3, ncol= 4, byrow= TRUE)
mat_1


mat_2= matrix(vec4, nrow= 3, ncol= 4, byrow= FALSE)
mat_2
```
10.) Show the code you used to create "mat_1".

-mat_1= matrix(vec4, nrow= 3, ncol= 4, byrow= TRUE)

11.) Show the code you used to create "mat_2".

-mat_2= matrix(vec4, nrow= 3, ncol= 4, byrow= FALSE)
    
    
```{r}
x= c(5.2)
y= c("five point two")
z= c(1:5)


my_list_1= list(two=x, one=y, three=z)
my_list_1

my_list_1[[3]]

my_list_1$one
my_list_1[[2]]
```
12.) Show the R code you used to create "my_list_1'.

-x= c(5.2)
y= c("five point two")
z= c(1:5)

my_list_1= list(two=x, one=y, three=z)



13.) Show the R code that would select third element of the list.

-my_list_1[[3]]
    
14.)Show the R code that selects the list element with the name “one”. Note: there are at least two ways to do this!

-my_list_1$one
      
OR
      
-my_list_1[[2]]


```{r}
my_vec = rep(1:3, 5)
my_vec


my_bool_vec = my_vec==3
my_bool_vec

data.frame(my_vec, my_bool_vec)


subset(my_vec, subset=my_bool_vec==TRUE)
```

15.) Show the R code that you used to create "my_bool_vec".

-my_bool_vec = my_vec==3
      
16.) Show the R code that you used to perform the subsetting.

-subset(my_vec, subset=my_bool_vec==TRUE)
      
      

