---
title: "In-Class Models 1"
author: "JT Larkin"
date: "11/2/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
require(here)
require(palmerpenguins)
```
```{r}
dat_ade = droplevels(subset(penguins, species == "Adelie"))

hist(dat_ade$body_mass_g, main = "Adelie Penguins: Body Mass", xlab = "body mass (g)")
```
Q1
```{r}
boxplot(body_mass_g ~ sex,
        data = dat_ade,
        xlab = "Sex",
        ylab = "Body Mass (g)",
        main = "Body Mass of Adelie Penguins")

```

Q2
```{r}
#?t.test()

dat_fem = droplevels(subset(dat_ade, sex == "female"))

t.test(dat_fem$body_mass_g, alternative = c("two.sided"), mu = 0)
```
Two-tailed t-test because it does not specify if its greater or less than zero grams. We just want to know if it is different than zero.


Q3

Yes, the mean is significantly different than zero. We are asserting the alternative that there is a difference. 


Q4
```{r}
dat_mal = droplevels(subset(dat_ade, sex == "male"))

t.test(dat_mal$body_mass_g, alternative = "greater", mu = 4000)
```
This is a one-tailed t-test. We specify that we want to know if the mean body mass is greater than 4000 g. 


Q5 
We are accepting the null that the mean body mass of male Adelie penguins is not greater than 4000. We would a mean body mass less than 4000 g approximately 1/15 experiments.

Q6

```{r}
t.test(dat_fem$body_mass_g, y = dat_mal$body_mass_g, alternative = "two.sided")
```

Q7

Yes, there is a significant difference between male and female body masses.



Q8

```{r}
t.test(dat_fem$body_mass_g, y = dat_mal$body_mass_g, alternative = "less")

# or 

t.test(dat_mal$body_mass_g, y = dat_fem$body_mass_g, alternative = "greater")
```

Q9
```{r}
t.test(dat_mal$body_mass_g, y = dat_fem$body_mass_g, alternative = "less")
```

Q10
Because practically all of the male mean weights will be greater than female mean weights, a p-value of one makes sense. Basically, its 1 - the p-value from question 8 which is practically 1.
