---
title: "Using Models 2"
author: 'JT Larkin (Partners: None)'
date: "11/20/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r include=FALSE}
require(here)
require(palmerpenguins)

t.test(subset(penguins, species == "Gentoo")$flipper_length_mm)

t.test(
x = subset(penguins, species == "Gentoo")$flipper_length_mm,
mu = 218
)

t.test(
x = subset(penguins, species == "Gentoo")$flipper_length_mm,
mu = 218,
alternative = "less"
)
```
```{r include=FALSE}
#Two-sample t.test
t.test(flipper_length_mm ~ species,
       data = subset(penguins, species != "Chinstrap"))

#1-way ANOVA
##Graphical
par(mfrow = c(1, 2))
hist(penguins$body_mass_g, breaks = 80, main = "histogram of body mass",
xlab = "body mass (g)")

plot(density(penguins$body_mass_g, na.rm = TRUE), 
     main = "density plot of body mass")

require(palmerpenguins)
boxplot(body_mass_g ~ species, data = penguins)

##Numerical
dat_chinstrap = subset(penguins, species == "Chinstrap")
mean(dat_chinstrap$body_mass_g, na.rm = TRUE)

shapiro.test(dat_chinstrap$body_mass_g)

aggregate(body_mass_g ~ species, data = penguins, FUN = mean)

##Fit Linear Model
fit_species = lm(body_mass_g ~ species, data = penguins)
summary(fit_species)

```


```{r include=FALSE}
#Conduct the ANOVA
anova(fit_species)

#One-Way Anova Complete Walkthrough
fit_species = lm(body_mass_g ~ species, data = penguins)
summary(fit_species)

```

```{r include=FALSE}
#Two-Way Additive ANOVA

boxplot(body_mass_g ~ species, data = penguins)

fit_additive = lm(body_mass_g ~ sex + species, data = penguins)
fit_additive

fit_interactive = lm(body_mass_g ~ sex * species, data = penguins)
fit_interactive
summary(fit_interactive)
```

```{r include=FALSE}
#Simple Linear Regression: Penguin Bills and Body Mass

require(palmerpenguins)

lm(bill_length_mm ~ body_mass_g, data = penguins)
```

################################################################################
Assignment


1. 
```{r}
boxplot(body_mass_g ~ sex * species, data = penguins,
    main = "Boxplot Of Penguin Body Mass Conditioned \nOn Both Sex And Species",
        xlab = "",
        ylab = "Body Mass (g)",
        las = 2,
        names = c("Female \nAdelie", "Male \nAdelie", "Female \nChinstrap",
                  "Male \nChinstrap", "Female \nGentoo", "Male \nGentoo"))
```


2. No. I came to this conclusion because the error bars for males and females overlap for all 3 species. If the error bars did not overlap I would say that there was a significant difference in mass.


3. Yes. When only species in the model the boxplot suggests that there may be an issue meeting the homogeneity assumption (some boxes widths are larger, thus more variance in some species means compared to others). Adding sex to the model makes the variances more uniform (in comparison to species alone). Thus, adding sex will help us meet the homogeneity assumption.

4. 

```{r}
fit_both = lm(body_mass_g ~ sex * species, data = penguins)
fit_both
summary(fit_both)
```
5. "sexfemale" is the base case for the two-way model.


6. "sexfemale" and "specieschinstrap"

7.  3527.21 grams
```{r}
3368.84 + 158.37
```


8. 3527.206 grams
```{r}
penguins = penguins
fem_chin = aggregate(
  body_mass_g ~ species * sex,
  data = penguins, FUN = mean)

fem_chin
```


